---
title: "Your Title"
author: "Your Name"
date: "Year Month Day"
output: html_document
---

This is a _small_ or _minimal_ R Markdown template for our course.  
A more extensive introduction is available [here](http://Rpository.com/down/temp/Rmarkdown_basics.html)  
(and the corresponding `.Rmd` source [here](http://Rpository.com/down/temp/Rmarkdown_basics.Rmd)).

# Load packages

```{r load_packages, echo = FALSE, eval = TRUE}
library("rmarkdown", "knitr")
```

# Text vs. Code

Write your text here. 

Put your R code in chunks:

```{r chunk_name}
# A unique chunk_name helps to keep track of your various chunks later. 
# R will evaluate the code in this chunk and print the result:
v <- 1:4
sum(v)
```

# Knitting

To convert an `.Rmd` source file into `.html` or `.pdf` output you need to "knit" it (`Command + Shift + K`). 

If your R code contains errors, the document will fail to knit (and print an error message that specifies the reason or chunk name with problems). Thus, if your document fails to knit, you need to go through your code (in a chunk-by-chunk fashion) to find and fix the error.

See [http://rmarkdown.rstudio.com](http://rmarkdown.rstudio.com) for further information.

<!-- End of document. -->