---
title: "ds4psy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ds4psy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r preamble_ds4psy, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- ds4psy logo 1: -->
<a href="https://bookdown.org/hneth/ds4psy/">
<img src = "./../inst/images/ds4psy.png" alt = "ds4psy" width = "125px" align = "right" style = "width: 125px; float: right; border:11;"/>
</a>

**Welcome** to the R package **ds4psy** --- a companion to the book and course  
[Data Science for Psychologists](https://bookdown.org/hneth/ds4psy/).

The course introduces the principles and methods of data science for students of psychology and other biological or social sciences. 


## Installation

The current release of **ds4psy** is available from [CRAN](https://CRAN.R-project.org/) at <https://CRAN.R-project.org/package=ds4psy>:

```{r install_CRAN, echo = TRUE, eval = FALSE}
install.packages('ds4psy')  # install ds4psy from CRAN client
library('ds4psy')           # load to use the package
```

The current development version can be installed from its [GitHub](https://github.com) repository at <https://github.com/hneth/ds4psy/>: 

```{r install_github, echo = TRUE, eval = FALSE}
# install.packages('devtools')  # (if not installed yet)
devtools::install_github('hneth/ds4psy')
library('ds4psy')  # load to use the package
```


## Resources

This package and the corresponding book are still being developed and are updated as new materials become available.  

- A current version of the book is available at <https://bookdown.org/hneth/ds4psy/>.  

- There are 2 GitHub repositories to be distinguished:  

    - The repository for the [ds4psy book](https://bookdown.org/hneth/ds4psy/) is <https://github.com/hneth/ds4psy_book>.  

    - The repository for the [ds4psy package](https://github.com/hneth/ds4psy/) is <https://github.com/hneth/ds4psy>.  

- The course syllabus (for Spring 2019) and all data files used in examples and exercises are available at <http://rpository.com/ds4psy/>.   


<!-- Update: -->

[Updated `r format(Sys.time(), "%Y-%m-%d")` by [hn](https://neth.de).]  

<!-- eof. -->